Optimizing Structural Properties Of Neural Networks With Genetic Algorithms

### 2015-08-06 paper/paper.tex
introduction adapted

    \usepackage{microtype}
    \usepackage{blindtext}
    \usepackage{enumitem}

@@

\section{Introduction}
Looking at the field of artificial neuronal networks (NN), one finds a great
variety of different models describing both the dynamic of networks as well
as their plasticities, i.e.\ their learning behavior.
But what exactly makes artificial neuronal networks learn successfully? And
why are certain types of networks and certain learning rules effective for
some problems, but fail at solving others? Confronting and eventually
answering these questions is critical when trying to gain a deep
understanding of neural networks and artificial intelligence.

Using the above questions as a guideline, we decided to analyze what
network characteristics make the FORCE algorithm \citep{FORCE} especially suitable
for learning periodic patterns.  More
precisely, we wanted to identify which properties of a NN are
important for its learning success, and how to choose their
values for an optimal result.  To this end, one option would be to take an
initial network and an adequate periodic pattern to be learned, and
optimize the network's internals for this very task. Then patterns may
emerge that hint at how an optimal network design looks.

However, we took a different approach:
Instead of fine tuning single internal weights, we were concerned with the
general rules of the network's structure. So we wanted our networks to
be random to a certain degree, which seems biologically plausible, but we
also wanted them to follow certain patterns and allow for specialization:
How densely are the synapses distributed? Is there a strong compartmentalization?
How strong should the feedback signal
be, and how strong should the synapses fire on average?

To address these issues we looked at parameterized random networks (in
the language of genetics, as we are using a genetic optimization algorithm, at different 
\emph{network genotypes}). These
genotypes contain parametric information about the networks
(\emph{phenotypes}) to be created. 
Different phenotypes of the same genotype may thus have strong
quantitative differences, but they share structural qualities whose fitness
for a preassigned range of tasks we wanted to analyze and improve.

% TODO we say the same thing later on again. maybe, the comment about the genetic algorithm
% should be moved out of the parenthesis
%Since calculating the fitness of a genotype requires the evaluation of many
%different phenotypes and because of the resulting stochastic nature of the
%problems we faced, we decided to use genetic algorithms for the
%optimization process.

>>>>

    \usepackage{microtype}
    \usepackage{blindtext}
    \usepackage{enumitem}
    \usepackage{siunitx}

@@

\section{Introduction}
Looking at the field of artificial neuronal networks (NN), one finds a great
variety of different models describing both the dynamic of networks as well
as their plasticities, i.e.\ their learning behavior. 
But, given some dynamics and plasticity rules, which features of a network
determine, if it is a good candidate for doing a specific task?
%But what exactly makes artificial neuronal networks learn successfully? And
%why are certain types of networks and certain learning rules effective for
%some problems, but fail at solving others? 
%Confronting and eventually
%answering these questions is critical when trying to gain a deep
%understanding of neural networks and artificial intelligence.

Using this question as a guideline, we decided to examine network
characteristics that make the FORCE algorithm \citep{FORCE} especially
suitable for learning periodic patterns. More precisely, we wanted to
identify which high level properties of a NN are important for its learning
success under FORCE, and how to choose their values for an optimal result.

%To this end, one option would be to take an
%initial network and an adequate periodic pattern to be learned, and
%optimize the network's internals for this very task. Then patterns may
%emerge that hint at how an optimal network design looks. 
%However, we took a different approach:
Instead of fine tuning single internal weights, we were thus concerned with the
general rules of the network's properties. And we wanted our networks to
be random to a certain degree, which seems biologically plausible; but we
also wanted to allow for structural specialization:
How densely are the synapses distributed? How strong shall the feedback
signal be, and how strong should the synapses fire on average?
%To address these issues we looked at parameterized random networks (in
%the language of genetics, as we are using a genetic optimization algorithm, at different 
%\emph{network genotypes}).

To address these issues we introduce \emph{network genotypes}, which
contain parametric information about the networks (\emph{phenotypes}) to be
created: Different phenotypes of the same genotype may have strong
differences in detail, but they share structural qualities of interest.
Improving these genotypes for a specific range of tasks via genetic
optimization is the main concern of this article.
%The fitness of these genotypes for a specific range of tasks was optimized 
%using genetic algorithms.
%whose fitness
%for a preassigned range of tasks we analyze and improve with genetic
%optimization.
% TODO we say the same thing later on again. maybe, the comment about the genetic algorithm
% should be moved out of the parenthesis
%Since calculating the fitness of a genotype requires the evaluation of many
%different phenotypes and because of the resulting stochastic nature of the
%problems we faced, we decided to use genetic algorithms for the
%optimization process.
